---
title: "CS273Project"
output: html_document
root.dir: "~/Google Drive/uci/2019Fall/CS273A/Project/dataset_diabetes"
editor_options: 
  chunk_output_type: console
---

```{r}
library(foreign)
library(dplyr)
library(tidyr)
#library(lattice)
library(ggplot2)
library(tidyverse)
library(lme4)
library(merTools)
#library(geepack)
#library(emdbook)
library(knitr)
library(plotROC)
library(stringr)
library(factoextra)
library(FactoMineR)
#for multinomial regression
library(nnet)
library(dummies)
library(glmnet)
#source("Allfunctions.R")

#load in balloonPlot
```
Diabetes or not relating to hospital readmission rate, first convert three diagnoses into coding of disease
```{r}
#read in data
diab = read.csv("~/Google Drive/uci/2019Fall/CS273A/Project/dataset_diabetes/diabetic_data.csv",header= T)
#ID = read.csv("~/Google Drive/uci/2019Fall/CS273A/Project/dataset_diabetes/IDs_mapping.csv",header = T)

#diab_JW = read.table("~/Google #Drive/uci/2019Fall/CS273A/Project/dataset_diabetes/hospital0.2.txt",header = T,fill = T)
diab2 = diab 
```
mutate disease into categories 
convert numeric into numeric
```{r}
diab2 = as.data.frame(apply(diab2,2,function(x) na_if(x, "?")))
rm(diab_JW)
#tmp  = diab2[,1:18] %>% select_if(~ nlevels(.) > 20 ) %>% mutate_if(is.factor, as.numeric)
#don't know how to put it back so going to do that one by one 

tmp = diab2 %>% mutate( encounter_id = as.numeric(encounter_id),patient_nbr = as.numeric(patient_nbr),time_in_hospital = as.numeric(time_in_hospital),num_lab_procedures = as.numeric(num_lab_procedures),num_procedures = as.numeric(num_procedures),num_medications = as.numeric(num_medications),number_outpatient = as.numeric(number_outpatient),number_emergency = as.numeric(number_emergency),number_inpatient = as.numeric(number_inpatient),number_diagnoses = as.numeric(number_diagnoses))

drop_cols = c("weight","payer_code","examide","citoglipton","medical_specialty")
tmp2 = tmp %>% dplyr::select(-one_of(drop_cols))
tmp3 = tmp2 %>% mutate(diag1 = as.character(diag_1),diag2 = as.character(diag_2),diag3 = as.character(diag_3))
```
check diagnosis and convert to dianose names 

```{r}
fit = lm(readmitted ~.,data = tmp2[,3:ncol(tmp2)],na.action = NULL )

```


```{r}
#this is for finding out what's in the diag
#grep("9999",tmp3$diag1)

#because in diagnoses there's letter, replace letter with 9999 for later mapping
tmp3$diag1 = as.numeric(gsub(x= as.character(tmp3$diag_1), pattern = "[A-z]\\d+", replacement = "9999"))
tmp3$diag2 = as.numeric(gsub(x= as.character(tmp3$diag_2), pattern = "[A-z]\\d+", replacement = "9999"))
tmp3$diag3 = as.numeric(gsub(x= as.character(tmp3$diag_3), pattern = "[A-z]\\d+", replacement = "9999"))

tmp3 = tmp3 %>% mutate(diagnose1 = case_when( (diag1 >= 390 & diag1<= 459) | diag1 ==785 ~ "Circulatory", (diag1 >= 460 & diag1<= 519) | diag1 ==786 ~ "Respiratory",(diag1 >= 520 & diag1<= 579) | diag1 ==787 ~ "Digestive", (diag1 > 249.00 & diag1 <251.00) ~ "Diabetes", (diag1 >= 800 & diag1<= 999) ~ "Injury",(diag1 >= 710 & diag1 <= 739) ~ "Musculoskeletal",(diag1 >= 580 & diag1<= 629) | diag1 ==788 ~ "Genitourinary",(diag1 >= 140 & diag1<= 239) | diag1 ==780 | diag1 ==781 |diag1 ==784 | (diag1 >= 790 & diag1 <= 799) | ( (diag1 >= 240 & diag1<= 279) & !(diag1 > 249.00 & diag1 <251.00)) | (diag1 >= 680 & diag1 <= 709)| diag1 ==782  |(diag1 >= 001 & diag1<=139)  ~ "Neoplasms",(diag1 >= 290 & diag1 <= 319)| (diag1 == 9999) | (diag1 >= 280 & diag1 <= 289) | (diag1 >= 320 & diag1 <= 359)| (diag1 >= 630 & diag1 <= 679)| (diag1 >= 360 & diag1 <= 389)|  (diag1 >= 740 & diag1 <= 759) | diag1 == 783 | diag1 == 789~ "Other")) %>%
  mutate(diagnose2 = case_when( (diag2 >= 390 & diag2<= 459) | diag2 ==785 ~ "Circulatory", (diag2 >= 460 & diag2<= 519) | diag2 ==786 ~ "Respiratory",(diag2 >= 520 & diag2<= 579) | diag2 ==787 ~ "Digestive", (diag2 > 249.00 & diag2 <251.00) ~ "Diabetes", (diag2 >= 800 & diag2<= 999) ~ "Injury",(diag2 >= 710 & diag2 <= 739) ~ "Musculoskeletal",(diag2 >= 580 & diag2<= 629) | diag2 ==788 ~ "Genitourinary",(diag2 >= 140 & diag2<= 239) | diag2 ==780 | diag2 ==781 |diag2 ==784 | (diag2 >= 790 & diag2 <= 799) | ( (diag2 >= 240 & diag2<= 279) & !(diag2 > 249.00 & diag2 <251.00)) | (diag2 >= 680 & diag2 <= 709)| diag2 ==782  |(diag2 >= 001 & diag2<=139)  ~ "Neoplasms",(diag2 >= 290 & diag2 <= 319)| (diag2 == 9999)| (diag2 >= 280 & diag2 <= 289) | (diag2 >= 320 & diag2 <= 359)| (diag2 >= 630 & diag2 <= 679)| (diag2 >= 360 & diag2 <= 389)|  (diag2 >= 740 & diag2 <= 759) | diag2 == 783 | diag2 == 789 ~ "Other")) %>% 
  mutate(diagnose3 = case_when( (diag3 >= 390 & diag3<= 459) | diag3 ==785 ~ "Circulatory", (diag3 >= 460 & diag3<= 519) | diag3 ==786 ~ "Respiratory",(diag3 >= 520 & diag3<= 579) | diag3 ==787 ~ "Digestive", (diag3 > 249.00 & diag3 <251.00) ~ "Diabetes", (diag3 >= 800 & diag3<= 999) ~ "Injury",(diag3 >= 710 & diag3 <= 739) ~ "Musculoskeletal",(diag3 >= 580 & diag3<= 629) | diag3 ==788 ~ "Genitourinary",(diag3 >= 140 & diag3<= 239) | diag3 ==780 | diag3 ==781 |diag3 ==784 | (diag3 >= 790 & diag3 <= 799) | ( (diag3 >= 240 & diag3<= 279) & !(diag3 > 249.00 & diag3 <251.00)) | (diag3 >= 680 & diag3 <= 709)| diag3 ==782  |(diag3 >= 001 & diag3<=139)  ~ "Neoplasms",(diag3 >= 290 & diag3 <= 319)| (diag3 == 9999)| (diag3 >= 280 & diag3 <= 289) | (diag3 >= 320 & diag3 <= 359)| (diag3 >= 630 & diag3 <= 679)| (diag3 >= 360 & diag3 <= 389)|  (diag3 >= 740 & diag3 <= 759) | diag3 == 783 | diag3 == 789~ "Other"))

#choose first visit 
#only keeping first visit 
tmp4 = as.data.frame(tmp3 %>% group_by(patient_nbr,.drop=T) %>%
  arrange(encounter_id) %>%
  filter(row_number()==1),.preserve = F)
```
1. save clean output as character
2. save the boolean vector matrix for python input 
```{r}
```
load in priyam's file and see if I can work on it, it works. not as .txt but as .csv, yes I can. 

Only choose subsets to do linear regression on, and also change diagnoses to. 
```{r}
drop_cols = c("glimepiride.pioglitazone")
datain = tmp4[,-c(16,17,18,46,47,48)] %>% dplyr::select(-one_of(drop_cols))
```
Analysis starts from here 
data called tmp4 
there are factor levels with 0 in one level, possibly due to choosing only first visits, needs to delte them Checked and it's true. 
```{r}

drop_cols = c("diag_1","diag_2","diag_3","diag1","diag2","diag3","glimepiride.pioglitazone","acetohexamide","metformin.pioglitazone","metformin.rosiglitazone")
tmp4 = tmp4 %>% dplyr::select(-one_of(drop_cols)) %>% 
  mutate_if(is.factor,droplevels)  %>%
  mutate(diagnose1 = as.factor(diagnose1),diagnose2 = as.factor(diagnose2),diagnose3 = as.factor(diagnose3))

#write_csv(tmp4 %>% mutate_if(is.factor,as.character),'Diabetes_clean.csv')

#doesn't work 
#datain %>% select_if(is.factor) %>% select_if(. %in% (levels(.)[summary(.) %in% 1])) %>% glimpse
# ind = rep(0,ncol(datain))
# for (i in 1:ncol(datain)) {
#    if (is.factor(datain[,i])) {
#      print(paste("here",as.character(i)))
#     ind[i] = sum(datain[,i] %in% levels(datain[,i])[summary(datain[,i]) %in% 1])
#     print(as.character(summary(datain[,i]) %in% 1))
#   }
# }
#doen checking, no more factor level less than 1 don't know why. gave up today 

datain = tmp4 %>% 
    dplyr::filter(!(miglitol %in% (levels(miglitol)[summary(miglitol) %in% 1]))) %>% 
    dplyr::filter(!(chlorpropamide %in% (levels(chlorpropamide)[summary(chlorpropamide) %in% 1]))) %>%
    dplyr::filter(!(discharge_disposition_id %in% (levels(discharge_disposition_id)[summary(discharge_disposition_id) %in% 1]))) %>%
    dplyr::filter(!(admission_source_id %in% (levels(admission_source_id)[summary(admission_source_id) %in% 1]))) %>% 
    mutate_at(c("miglitol","chlorpropamide","admission_source_id","discharge_disposition_id"),droplevels)
datain = na.omit(datain) %>% 
  dplyr::filter(!(gender %in% (levels(gender)[summary(gender) %in% 1]))) %>% 
  mutate_at(c("gender"),droplevels) 
load("FinalData.Rdata")
#load in datain 

#doesn't support filter for summary for the column instead of column names 
#dplyr::filter_at(vars(any(summary(.) %in% 1)),!(. %in% (levels(.)[summary(.) %in% 1])))
```
###################### Analysis starts from here 
PCA analysis 
```{r}
#convert everything except for response variable into dummy variable for easier time of everything
#load in txt 
data2=read_csv("hospital_ready.csv")

#merge readmitted >30 and <30 into yes or no 
datain= datain %>% mutate(readmit2= case_when(readmitted == "NO" ~"NO", readmitted == ">30"| readmitted == "<30" ~ "YES")) %>% 
  mutate(readmit2 = as.factor(readmit2)) %>% 
  mutate(discharge_disposition_id= as.numeric(discharge_disposition_id)) %>% 
  mutate(admission_source_id=as.numeric(admission_source_id),admission_type_id = as.numeric(admission_type_id)) %>% 
  mutate(age = as.numeric(age))
datain2 = dummy.data.frame(datain,names = c("race","diagnose1","diagnose2","diagnose3","change","diagbetesMed","gender"))
datain2 = datain2 %>% mutate_if(is.factor,as.numeric)

dataplot = datain %>% mutate(discharge_disposition_id = as.numeric(discharge_disposition_id))

#this replicate Jinwei's remapping 
save(datain2,file="FinalDataDummy.RData")

#these two libraries for plotting PCA
data2=read_csv("hospital_ready.csv")

outpca = prcomp(datain2 %>% dplyr::select(-readmit2,-readmitted,-encounter_id,-patient_nbr),scale.=T)
outpca = prcomp(data2[,-ncol(data2)],scale.=T)
old.par <- par(mar = c(0, 0, 0, 0))
par(old.par)
fviz_pca_var(outpca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
# res = MFA(tmp4[,-c(1,2,39)], group=c(3,3,3,10,9,3), type=c("n",rep("s",5)), ncp=5, name.group=c("ind","hospital","visual","odor.after.shaking", "taste","overall","diagnoses"), num.group.sup=c(1,6))
```
########Outliers and plotting 
Now using only complete dataset and deletes variable payer_code and because there is 53% missing.  
```{r}
#simple linear regression 
#see if it will run after deleting those NA rows, possibly due to less factor levels after removing NAs
#delete medical specialty and payer code because too much NA causing trouble for other variables to miss factors levels as well
datain[,-c(1,2)] %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density()  


datain[,-c(1,2)] %>%
  keep(is.factor) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_bar()  

#try
mtcars %>%
  gather(-mpg, -hp, -cyl, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y = mpg, color = hp, shape = factor(cyl))) +
    geom_point() +
    facet_wrap(~ var, scales = "free") +
    theme_bw()

out2 = glm(as.factor(readmit2) ~ .,data = datain2[,-c(1,2,44)],family=binomial)
out3 = multinom(as.factor(readmitted) ~ .,data = datain2[,-c(1,2,ncol(datain2))])
out = lm(readmitted~ .,data = datain2[,-c(1,2,ncol(datain2))])
#already have significant results 
```
########
best subsets
multinomial regression and best subsets 
use AIC as a criterion 
find best subsets 

1. test if predicted needs to also be factor 
```{r}
dataUse =  datain2[,-c(1,2,ncol(datain2))]
dataUse$readmitted = as.factor(dataUse$readmitted)

#write code to divide data into 10 fold 
#don't like it try the glm binary first 
# alpha=0.7
# d = sort(sample(nrow(dataUse), nrow(dataUse)*alpha))
# train = datain[d,]
# test = dataUse[-d,]
# glm.fit=multinom(as.factor(readmitted)~., data=data2)
# predict(glm.fit, test, "probs")
# mean(as.character(predict (glm.fit, test[,-ncol(test)])) != as.character(test$cmc))

#use logistic best subsets for speed and timing, use the same model for final prediction 
#library(bestglm)
dataUseglm = datain2[,-c(1,2,44)] %>% mutate(y=as.factor(readmit2)) %>% dplyr::select(-readmit2)
#doest work
#res.bestglm = bestglm(Xy = dataUseglm,family=binomial,IC="AIC",method="exhaustive")

dataUseglm2 = datain2[,-c(1,2,44)] %>% dplyr::select(-raceOther,-genderMale,-changeNo, -diagnose1Respiratory,-diagnose2Respiratory,-diagnose3Respiratory)
#outglmlsa = glm(as.factor(readmit2) ~ .,data = dataUseglm2,family=binomial)
#lsa(outglmlsa) #does give meaningful output for variable 


y <- as.numeric(datain2$readmitted)
x=as.matrix(dataUseglm2[,-c(ncol(dataUseglm2))])

#save the prediction output 
#also code for plotting AUC curve 
#try gaussian, if not long, then use multinom  2

cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "multinomial")
plot(cv.lasso)
Accuracy = rep(0,10)
for (i in 1:10) {
alpha=0.7
d = sort(sample(nrow(dataUseglm2), nrow(dataUseglm2)*alpha))

Xtrain = as.matrix(dataUseglm2[d,-ncol(dataUseglm2)])
Xtest = as.matrix(dataUseglm2[-d,-ncol(dataUseglm2)])
Ytrain = as.numeric(datain2$readmitted)[d]
Ytest = as.numeric(datain2$readmitted)[-d]

model <- glmnet(x, y, alpha = 1, family = "multinomial",
                lambda = cv.lasso$lambda.min)
#test if can get prediction 1 works

predictedclass <- model %>% predict(newx = Xtest,typ="class")
Accuracy[i]= mean(as.numeric(predictedclass) == Ytest)

}


save(dataUseglm2,file="dataUseGLM2.RData")
save(datain,file="datain.RData")
save(dataUseglm,file="dataUseGLM.RData")
#output best subsets 
```
simpler way to do cross validation 
```{r}
#with chosen alpha, now do 
```
